---
title: "create sample name file for dadasnake"
output: html_document
date: "2023-10-26"
---

load libraries
```{r}
library(utils)
library(purrr)
```

first, unzip folder with fastq reads on Y drive. then, set working directory to folder containing unzipped fastq files
```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = "Y:/ABL_Genetics/Lab_data/MiSeq_data/kim_temporary/20240820_nsmock/")
```


get sample names and r1 and r2 file names to use with dadasnake 
```{r}
# List the files with a .fastq extension
files <- list.files(pattern = "\\.fastq.gz$")

# Filter out files that start with "Undetermined"
files <- files[!grepl("^Undetermined", files)]

## if multiple projects/amplicons were lumped onto the same miseq run, divide up into in folders.  
#files <- files[!grepl("18S", files)]
#files <- files[!grepl("MiFish", files)]

# Initialize data frame with a single row
df <- data.frame(
  sample = character(1),
  library = character(1),
  run = numeric(1),
  r1_file = character(1),
  r2_file = character(1),
  stringsAsFactors = FALSE
)

# Initialize vectors for storing sample IDs
sample_ids <- character(0)

# Iterate through the remaining files
for (filename in files) {
  parts <- unlist(strsplit(filename, "_"))
  if (length(parts) >= 3) {
  #if (length(parts) >= 1) {
    library <- parts[1]
    is_r1 <- grepl("_R1_", filename)
    is_r2 <- grepl("_R2_", filename)
    #is_r1 <- grepl("_R1.", filename)
    #is_r2 <- grepl("_R2.", filename)

    if (library %in% sample_ids) {
      # Update existing row
      idx <- which(df$sample == library)
      if (is_r1) {
        df[idx, "r1_file"] <- filename
      }
      if (is_r2) {
        df[idx, "r2_file"] <- filename
      }
    } else {
      # Add a new row
      sample_ids <- c(sample_ids, library)
      new_row <- data.frame(
        sample = library,
        library = library,
        run = 2,
        r1_file = ifelse(is_r1, filename, ""),
        r2_file = ifelse(is_r2, filename, ""),
        stringsAsFactors = FALSE
      )
      df <- rbind(df, new_row)
    }
  }
}

# Remove the initial empty row - if combining multiple miseq runs, each will need to be processed seperately
#df_run1 <- df[-1, ]
#df_run2 <- df[-1, ]
#df_run3 <- df[-1, ]
#df_run4 <- df[-1, ]

df <- df[-1, ]
```

combine multiple runs, if applicable
```{r}
# library(dplyr)
# df_all <- df_run1 %>%
#   bind_rows(df_run2) %>%
#   bind_rows(df_run3) %>%
#   bind_rows(df_run4)
```

add seq date to sample/library name so that dadasnake keeps the twiced sequenced samples separate, if applicable 
```{r}
# df_all_w_date <- df_all %>%
#   mutate(date = ifelse(run == "1", "20240423", NA),
#          date = ifelse(run == "2", "20240509", date),
#          date = ifelse(run == "3", "20240611", date),
#          date = ifelse(run == "4", "20240613", date)) %>%
#   unite(col = "sample_date", sample, date, sep = "-") %>%
#   mutate(library = sample_date) %>%
#   rename(sample = sample_date)
```


```{r}
# Write the data frame to a TSV file
write.table(df, file = "~/nsmock_samples.tsv", sep = "\t", quote = FALSE, row.names = FALSE)
```

upload .tsv to Sedna so it is within the same folder as the raw fastq files on Sedna 

remove unzipped folder from genetics shared drive after creating this file 
